Approximation as Differentiation

### 2025-06-17 main/higher-order.tex
Better to define 'small product' rather than 'category with all small products'.

$\Fam(\FinVect)$ and $\Fam(\LatGal)$, introduced in \secref{first-order}, provide appropriate notions of
composition for automatic differentiation and Galois slicing respectively, but neither is Cartesian closed and
thereby able to interpret higher-order programs directly. However the following is a special case of a
construction due to \citet{nunes2023}:

\begin{proposition}
Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
$\Fam(\cat{C})$ is Cartesian closed.
\end{proposition}

A category $\cat{C}$ has all small (set-indexed) products if, for any set $X$ and any $X$-indexed family $A$
of objects of $\cat{C}$, there exists an object $\prod_{x \in X}A_x$ (also written $\prod_{X} A$) and family
of projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$
and family of morphisms $\{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
$\prodFam{f}$ making the following diagram commute for every $x \in X$:

\begin{center}
\begin{tikzcd}
   \prod_{X} A \arrow[r, "\pi_x"] & A_x \\
   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\prodFam{f}"]
\end{tikzcd}
\end{center}

>>>>

$\Fam(\FinVect)$ and $\Fam(\LatGal)$, introduced in \secref{first-order}, provide appropriate notions of
composition for automatic differentiation and Galois slicing respectively, but neither is Cartesian closed. To
interpret higher-order programs directly, we use a special case of a construction due to \citet{nunes2023}
(\propref{higher-order:fam-CC} below). First we introduce notation for the product over a set-indexed family.

\begin{definition}[Small product]
Suppose $\cat{C}$ a category and $X$ a
set. A \emph{product} of the family $A = \{A_x\}_{x \in X}$ is an object $\prod_{X}A$ plus a family of
projection morphisms $\{\pi_x: \prod_{X} A \to A_x\}_{x \in X}$ in $\cat{C}$ such that for any object $B$ and
family of morphisms $f = \{f_x: B \to A_x\}_{x \in X}$ in $\cat{C}$, there exists a unique product morphism
$\prodFam{f}$ making the following diagram commute for every $x \in X$:

\begin{center}
\begin{tikzcd}
   \prod_{X} A \arrow[r, "\pi_x"] & A_x \\
   B \arrow[ru, "f_x"'] \arrow[u, dotted, "\prodFam{f}"]
\end{tikzcd}
\end{center}
\end{definition}

\begin{proposition}
\label{prop:higher-order:fam-CC}
Suppose $\cat{C}$ locally small. If $\cat{C}$ has binary biproducts and all small products, then
$\Fam(\cat{C})$ is Cartesian closed.
\end{proposition}
